### 2021-08-10 Notes/notes.tex
created notes skeleton

\documentclass[onehalfspacing]{article}

\usepackage{amsmath,amssymb,amsthm}
\usepackage{ wasysym }

\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{definition}[theorem]{Definition}

\author{Alexander Pluska}
\title{Proofs as Programs in Classical Logic\\Notes}

\begin{document}

\maketitle

\bibliographystyle{acm}
\bibliography{references}

\end{document}