### 2022-07-07 CLS/lipics-v2021-EIICL.tex
CLS paper

\documentclass[a4paper,UKenglish,cleveref, autoref, thm-restate]{lipics-v2021}
%This is a template for producing LIPIcs articles. 
%See lipics-v2021-authors-guidelines.pdf for further information.
%for A4 paper format use option "a4paper", for US-letter use option "letterpaper"
%for british hyphenation rules use option "UKenglish", for american hyphenation rules use option "USenglish"
%for section-numbered lemmas etc., use "numberwithinsect"
%for enabling cleveref support, use "cleveref"
%for enabling autoref support, use "autoref"
%for anonymousing the authors (e.g. for double-blind review), add "anonymous"
%for enabling thm-restate support, use "thm-restate"
%for enabling a two-column layout for the author/affilation part (only applicable for > 6 authors), use "authorcolumns"
%for producing a PDF according the PDF/A standard, add "pdfa"

%\pdfoutput=1 %uncomment to ensure pdflatex processing (mandatatory e.g. to submit to arXiv)
%\hideLIPIcs  %uncomment to remove references to LIPIcs series (logo, DOI, ...), e.g. when preparing a pre-final version to be uploaded to arXiv or another public repository

\usepackage{tikz-qtree}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows.meta}
\usetikzlibrary{arrows,shapes,quotes}

%\graphicspath{{./graphics/}}%helpful if your graphic files are in another directory

\bibliographystyle{plainurl}% the mandatory bibstyle

\title{Embedding intuitionistic into classical logic} %TODO Please add

%\titlerunning{Dummy short title} %TODO optional, please use if title is longer than one line

\author{Alexander {Pluska}}{Faculty of Mathematics, Universtät Wien, Austria}{e11941874@student.tuwien.ac.at}{https://orcid.org/my-orcid?orcid=0000-0002-7709-3335}{}%TODO mandatory, please use full name; only 1 author per \author macro; first two parameters are mandatory, other parameters can be empty. Please provide at least the name of the affiliation and the country. The full address is optional. Use additional curly braces to indicate the correct name splitting when the last name consists of multiple name parts.

\author{Florian Zuleger}{Institute of Logic and Computation, Technische Universität Wien, Austria}{florian.zuleger@tuwien.ac.at}{https://orcid.org/0000-0003-1468-8398}{}

\authorrunning{A. Pluska and F. Zuleger} %TODO mandatory. First: Use abbreviated first/middle names. Second (only in severe cases): Use first author plus 'et al.'

\Copyright{Jane Open Access and Joan R. Public} %TODO mandatory, please use full first names. LIPIcs license is "CC-BY";  http://creativecommons.org/licenses/by/3.0/

\ccsdesc[100]{\textcolor{red}{Theory of Computation Logic Constructive Mathematics}} %TODO mandatory: Please choose ACM 2012 classifications from https://dl.acm.org/ccs/ccs_flat.cfm

\keywords{Intuitionistic Logic, Automated Reasoning, Proof Theory} %TODO mandatory; please add comma-separated list of keywords

\category{} %optional, e.g. invited paper

\relatedversion{} %optional, e.g. full version hosted on arXiv, HAL, or other respository/website
%\relatedversiondetails[linktext={opt. text shown instead of the URL}, cite=DBLP:books/mk/GrayR93]{Classification (e.g. Full Version, Extended Version, Previous Version}{URL to related version} %linktext and cite are optional

%\supplement{}%optional, e.g. related research data, source code, ... hosted on a repository like zenodo, figshare, GitHub, ...
%\supplementdetails[linktext={opt. text shown instead of the URL}, cite=DBLP:books/mk/GrayR93, subcategory={Description, Subcategory}, swhid={Software Heritage Identifier}]{General Classification (e.g. Software, Dataset, Model, ...)}{URL to related version} %linktext, cite, and subcategory are optional

%\funding{(Optional) general funding statement \dots}%optional, to capture a funding statement, which applies to all authors. Please enter author specific funding statements as fifth argument of the \author macro.

%\nolinenumbers %uncomment to disable line numbering

%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\EventEditors{John Q. Open and Joan R. Access}
\EventNoEds{2}
\EventLongTitle{42nd Conference on Very Important Topics (CVIT 2016)}
\EventShortTitle{CVIT 2016}
\EventAcronym{CVIT}
\EventYear{2016}
\EventDate{December 24--27, 2016}
\EventLocation{Little Whinging, United Kingdom}
\EventLogo{}
\SeriesVolume{42}
\ArticleNo{23}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\maketitle

%TODO mandatory: add short abstract of the document
\begin{abstract}
The famous double negation translation~\cite{glivenko1929quelques, godel1933intuitionistischen} establishes an embedding from classical into intuitionistic logic. Curiously the reverse direction has not been covered in literature, perhaps because going from the more expressive logic to the less expressive one necessitates a much more involved encoding. We present an embedding from intuitionistic to classical logic, both in the propositional and first-order case, as well as an embedding of intuitionistic propositional logic into quantified boolean formulas. 
Furthermore in the propositional case we establish a parameter that bounds the proof complexity of intuitionistic formulas. In general validity of intuitionistic propositional logic is \verb+PSPACE+-complete~\cite{statman1979intuitionistic} but this gives us a finer-grained result. In this process we also present a Tseytin-like translation that preserves intuitionistic validity and reduces formulas to a normal form.
Finally in the first order-case we hope that our translation allows the use state-of-the-art theorem provers for classical first-order logic such as Vampire~\cite{kovacs2013first} for checking intuitionistic validity and program synthesis. 
\end{abstract}

\section{Introduction}

Most state-of-the-art first-order theorem provers leverage classical logic~\cite{kovacs2013first, schulz2002brainiac, korovin2008iprover}. Besides historic reasons a main advantage of classical logic over intuitionistic logic is the existence of convenient normal forms which enable reasoning techniques such as superposition and resolution. However there are also some drawbacks of this approach, in particular the non-constructiveness of classical logic poses a problem in synthesis-related tasks. On the other hand the Curry-Howard correspondence makes program extraction from intuitionistic proofs trivial~\cite{howard1980formulae} but automated deduction for intuitionistic logic is notoriously difficult.

There are different approaches for program extraction using classical provers. The first is simply to examine under which circumstances classical validity entails intuitionistic validity. Most famously this holds for $\Pi_2$-formulas in Peano Arithmetic~\cite{friedman1978classically}. There are also other circumstances under which this entailment holds~\cite{schwichtenberg}, however this approach cannot encapsulate the general case. Another popular approach is to equip classical proofs with operational semantics~\cite{Control1, Parigot1}. In this case we necessarily lose realizability at least of $\exists$ and $\vee$, which - one could argue - is why we want a program in the first place.

We examine a new direction, i.e. for each formula $\varphi$ giving a translation $\varphi^\#$ such that $\varphi$ is intuitionistically valid if and only if $\varphi^\#$ is classically valid. We shall start with the simpler case of propositional formulas. The process will be divided into three steps. The first is a first-order encoding of intuitionistic Kripke semantics, which gives an obvious but unsatisfactory embedding. We then discuss how some of the quantifiers can be eliminated via Herbrandization and give a Tseytin-like transformation that exposes a parameter bounding the proof complexity of the propositional formula. This finally allows us to eliminate all remaining quantifiers and we end up with a propositional formula. Since validity of intuitionistic logic is \verb+PSPACE+-complete while for classical logic it only \verb+coNP+-complete as one could expect there is a blow-up which is exponential in the parameter that we previously isolated.
Instead of targetting classical propositional logic at the last step we give an alternate translation to QBF, which is linear in the size of our complexity parameter. For predicate formulas the process is a bit more involved in that we also need to respect the constructive content of quantifiers. In particular the different semantics of the $\forall$-quantifier proves challenging. After dealing with this, we are able to utilize a quite similar process as in the propositional case.

\section{IPC to CQC}

The most obvious approach to embedding intuitionistic into classical logic is to express intuitionistic semantics, i.e. Kripke frames, as a first order theory. For every propositional variable $A$ consider a unary predicate $A$ of the same name where $A(u)$ expresses that $A$ is true at some world $u$. We can then naively encode formulas as follows:

\begin{definition}
	Let $\varphi$ be a propositional formula. Define $\varphi^{k}$ inductively as follows:
	\begin{itemize}
		\item $A^{k} = A(k)$ for every propositional variable $A$.
		\item $(\varphi\wedge\psi)^k = \varphi^k\wedge\psi^k$.
		\item $(\varphi\vee\psi)^k = \varphi^k\vee\psi^k$.
		\item $(\varphi\to \psi)^k = \forall u(k\preceq u\to\varphi^{u}\to\psi^{u})$ where $u$ is some new free variable.
	\end{itemize}
	Let $K(\varphi)$ encode that $\preceq$ is a partial order, the persistency condition for all propositional variables occurring in $\varphi$
	\begin{align*}
		K(\varphi) = &\forall k(k\preceq k)\wedge \forall k\forall l(k\preceq l\to l\preceq k\to l = k) \wedge \\&\forall k\forall l\forall m(k\preceq l \to l\preceq m\to k\preceq m)\wedge\\&\bigwedge \{\forall k\forall l(k\preceq l\to A(k)\to A(l))\:|\: \text{ $A$ occurs in $\varphi$}\}
	\end{align*}
	Then define
	$$\varphi^{C} = K(\varphi)\to \varphi^{u}$$
	where $u$ is some free variable.
\end{definition}

All that we have done is model Kripke frames with a finite number of propositional variables as a first-order theory and by definition we have

\begin{lemma}
	$\varphi$ is intuitionistically valid iff $\varphi^C$ is classically valid.
\end{lemma}

\begin{example}
	Consider the law of excluded middle $$\varphi = (A\to \bot)\vee A$$ which is not intuitionistically valid. Then
	\begin{align*}
		\varphi^{C} &= K(\varphi)\to ((A\to \bot)\vee A)^u\\
		&= K(\varphi)\to (A\to \bot)^u\vee A^u\\
		&= K(\varphi)\to (\forall u'( u\preceq u' \to A^{u'}\to \bot^{u'})\vee A(u))\\
		&= K(\varphi)\to (\forall u' (u\preceq u' \to A(u')\to \bot(u'))\vee A(u))
	\end{align*}
	and indeed $\varphi^{C}$ admits a counter model with domain $\{u, u'\}$ where $u \preceq u'$, $A(u) = 0, A(u') = 1$.
\end{example}

\section{Quantifiers in Classical Logic}

Of course translating a propositional formula to a first-order formula is somewhat unsatisfying. As a first step to obtaining a propositional formula we wish to eliminate the quantifiers. Since we are interested in preserving validity we use Herbrandization. In this process we introduce free variables and also additional function symbols to the signature. Whenever we speak of a new free variable what is meant is a free variable that does not occur in any of the considered formulas. Whenever a new function symbol is introduced we implicitly extend the signature by some not previously contained symbol.

Now usually in classical logic before removing quantifiers prenexification is performed, however since we wish to apply a similar procedure to intuitionistic logic later, where prenexification is not valid, we will do without.

\begin{definition}
	For formulas $\varphi$ we define the Skolemization $\varphi^S_T$ and Herbrandization $\varphi^H_T$ with respect to $T$ by simultaneous induction as follows:
	\begin{itemize}
		\item $A^S_T = A^H_T = A$ for each atomic $A$.
		\item $(\varphi\circ\psi)^X_T = \varphi^X_T\circ\psi^X_T$ for $\circ\in\{\wedge, \vee\}$, $X\in\{S, H\}$.
		\item $(\varphi\to\psi)^S_T = \varphi^H_T\to \psi^S_T$\\$(\varphi\to\psi)^H_T = \varphi^S_T\to\psi^H_T$.
		\item $(\forall x\varphi)^S_T = \forall x(\varphi[a/x]^S_{T\cup\{a\}}[x/a])$ where $a$ is a new free variable\\$(\forall x\varphi)^H_T = \varphi[s(t_1,\dots,t_n)/x]^H_T$ where $s$ is a new function, $\{t_1\dots t_n\} = T$.
		\item $(\exists x\varphi)^S_T = \varphi[s(t_1,\dots,t_n)/x]^S_T$ where $s$ is a new function, $\{t_1\dots t_n\} = T$\\$(\exists x\varphi)^H_T = \exists x(\varphi[a/x]^H_{T\cup\{a\}}[x/a])$ where $a$ is a new free variable.
	\end{itemize}
	Let $\varphi^S = (\exists x_1\dots\exists x_n \varphi[x_1/a_1\dots x_n/a_n])^S_\emptyset$ and $\varphi^H = (\exists x_1\dots\exists x_n \varphi[x_1/a_1\dots x_n/a_n])^H_\emptyset$ where $a_1,\dots,a_n$ are the free variables occurring in $\varphi$.
\end{definition}

\begin{theorem}For every formula $\varphi$
	\begin{itemize}
		\item $\varphi$ and $\varphi^S$ are classically equisatisfiable.
		\item $\varphi$ and $\varphi^H$ are classically equivalid.
	\end{itemize}
\end{theorem}

which follows from the following two Lemmata, proofs of which can be found in the appendix.

\begin{lemma}\label{ap1}
	Let $\chi$ be a formula and $\{t_1\dots t_n\} = T$ contain all free variables in $\chi$, then for every structure $\mathcal M = (M, I)$ there exists $\mathcal M_\chi = (M, I_\chi$) such that $p^{I_{\chi}} = p^I$ for all function and relation symbols $p$ occurring in $\chi$ and for all variable assignments $v$ we have
	\begin{itemize}
		\item if $\mathcal M, v\models\chi$ then $\mathcal M_\chi, v\models\chi^S_T$.
		\item if $\mathcal M, v\not\models\chi$ then $\mathcal M_\chi, v\not\models\chi^H_T$.
	\end{itemize}
\end{lemma}

\begin{lemma}\label{ap2}
	For every structure $\mathcal M$ and $\{t_1\dots t_n\} = T$ that contains all free variables in $\chi$ and variable assignment $v$
	\begin{itemize}
		\item if $\mathcal M, v\models\varphi^S_T$ then $\mathcal M, v\models \varphi$
		\item if $\mathcal M, v\not\models\varphi^H_T$ then $\mathcal M, v\not\models\varphi$.
	\end{itemize}
\end{lemma}

\section{Equivalid normalization}

Now trying to apply Herbrandization to arbitrary formulas presents us with quite a leap in complexity, leading to nested expressions that will make arguments more complicated. We therefore propose a Tseytin-like transformation that produces a formula of a more manageable syntactic form. A similar translation was proposed in~\cite{statman1979intuitionistic}.

\begin{definition}
	Let $\varphi$ be some formula. For each sub(semi)formula $\psi$ of $\varphi$ consider some new $n$-ary relation symbol $P_\psi$ where $n$ is the number of variables occurring in $\psi$ that is not quantified within $\psi$. We denote those variables with $\vec t_\psi$ in some arbitrary but fixed order. We inductively define clause sets $\mathcal S^+(\varphi)$ and $\mathcal S^-(\varphi)$ as follows:
	\begin{itemize}
		\item For every atomic formula $A$\\$\mathcal S^+(A) = \{P_A(\vec t_A)\to A\}$\\$\mathcal S^-(A) = \{A\to P_A(\vec t_A)\}$
		\item For every conjunction or disjunction $\varphi\circ\psi$ with $\circ\in\{\wedge,\vee\}$\\$\mathcal S^+(\varphi\circ\psi) = \{P_{\varphi\circ\psi}(\vec t_{\varphi\circ\psi})\to P_{\varphi}(\vec t_\varphi)\circ P_{\psi}(\vec t_\psi)\}\cup \mathcal S^+(\varphi)\cup \mathcal S^+(\psi)$\\$\mathcal S^-(\varphi\circ\psi) =\{P_{\varphi}(\vec t_\varphi)\circ P_{\psi}(\vec t_\psi)\to P_{\varphi\circ\psi}(\vec t_{\varphi\circ\psi})\}\cup \mathcal S^-(\varphi)\cup \mathcal S^-(\psi)$
		\item For every implication $\varphi \to\psi$\\$\mathcal S^+(\varphi\to\psi) = \{(P_{\varphi\to\psi}(\vec t_{\varphi\to\psi})\wedge P_{\varphi}(\vec t_\varphi))\to P_{\psi}(\vec t_\psi)\}\cup \mathcal S^-(\varphi)\cup \mathcal S^+(\psi)$\\$\mathcal S^-(\varphi\to\psi)  = \{(P_{\varphi}(\vec t_\varphi)\to P_{\psi}(\vec t_\psi))\to P_{\varphi\to\psi}(\vec t_{\varphi\to\psi})\}\cup \mathcal S^+(\varphi)\cup \mathcal S^-(\varphi)$
		\item For quantified formulas $Qx\varphi(x)$ with $Q\in \{\forall,\exists\}$\\$\mathcal S^+(Qx\varphi(x)) = \{P_{Qx\varphi}(\vec t_{Qx\varphi})\to QxP_{\varphi}(\vec t_{\varphi})\}\cup \{\forall x\psi\:|\:\psi\in\mathcal S^+(\varphi)\}$\\$\mathcal S^-(Qx\varphi(x))  = \{QxP_{\varphi}(\vec t_{\varphi})\to P_{Qx\varphi}(\vec t_{Qx\varphi})\}\cup \{\forall x\psi\:|\:\psi\in\mathcal S^-(\varphi)\}$
	\end{itemize}
	and have $\mathcal S(\varphi) = \mathcal S^+(\varphi)\cup\mathcal S^-(\varphi)$.
\end{definition}

\begin{lemma}The following are true for both classical and intuitionistic semantics:
	\begin{itemize}
		\item $\bigwedge S^+(\varphi)\wedge P_\varphi\models\varphi$. 
		\item $\bigwedge S^-(\varphi)\wedge \varphi\models P_\varphi$. 
	\end{itemize}
\end{lemma}

\begin{proof}
	The claims are verified in a straightforward manner by induction on the height of $\varphi$.
\end{proof}

\begin{example}
	Consider $\varphi = \neg\forall x A(x)\to \exists x\neg A(x)$. Then
	\begin{align*}
		\mathcal S^+(\varphi) = &\{(P_\varphi\wedge P_{\neg\forall x A(x)})\to P_{\exists x\neg A(x)}, (P_{\forall xA(x)}\to\bot)\to P_{\neg \forall xA(x)}\}\\&\cup\{\forall xP_{A(x)}(x)\to P_{\forall xA(x)}, \forall x(P_{A(x)}(x)\to A(x))\}\\&\cup\{P_{\exists x\neg A(x)}\to \exists xP_{\neg A(x)}(x), \forall x(P_{\neg A(x)}(x)\to P_{A(x)}(x)\to \bot)\}\\&\cup\{\forall x(A(x)\to P_{A(a)}(x))\}\\
		\mathcal S^-(\varphi) = &\{(P_{\neg\forall x A(x)}\to P_{\exists x\neg A(x)})\to P_\varphi,  (P_{\neg \forall xA(x)}\wedge P_{\forall xA(x)})\to\bot\}\\&\cup\{P_{\forall xA(x)}\to \forall xP_{A(x)}(x), \forall x(A(x)\to P_{A(x)}(x))\}\\&\cup\{\exists xP_{\neg A(x)}(x)\to P_{\exists x\neg A(x)}, \forall x((P_{A(x)}(x)\to \bot)\to P_{\neg A(x)}(x))\}\\&\cup\{\forall x(P_{A(a)}(x)\to A(x))\}
	\end{align*}
\end{example}

We can give a transformation of structures that corresponds to the above transformation of sentences.

\begin{definition}
	For every classical and intuitionistic structure $\mathcal M$ define a structure $\mathcal S(\mathcal M,\varphi)$ that agrees with $\mathcal M$ on everything except the interpretation(s) of atoms $P_\psi$. By slight abuse of notation we denote with $\vec t_\psi$ elements of the domain instead of variables. In the classical case have	
	$$P_{\varphi\wedge\psi}^I(\vec t_{\varphi\wedge\psi}) \Leftrightarrow P_{\varphi}^I(\vec t_\varphi)\wedge P_{\psi}^I(\vec t_\psi)\indent P_{\varphi\vee\psi}^I(\vec t_{\varphi\vee\psi}) \Leftrightarrow P_{\varphi}^I(\vec t_\varphi)\vee P_{\psi}^I((\vec t_\psi))$$$$ P_{\varphi\to\psi}^I(\vec t_{\varphi\to\psi}) \Leftrightarrow (\neg P_{\varphi}^I(\vec t_{\varphi}))\vee P_{\psi}^I(\vec t_{\psi})$$
	$$P_{\forall x\varphi}^I(\vec t_{\forall x\varphi}) \Leftrightarrow \{\vec t\:|\:\:P_{\varphi}^I(\vec t_\varphi) \text{ for all $x\in M$}\}$$$$ P_{\exists x\varphi}^I(\vec t_{\exists x\varphi}) \Leftrightarrow \{\vec t\:|\:\:P_{\varphi}^I(\vec t_\varphi) \text{ for some $x\in M$}\}$$
	\\

	and for intuitionistic logic and some world $u$ 		
	$$P_{\varphi\wedge\psi}^{I_u}(\vec t_{\varphi\wedge\psi}) \Leftrightarrow P_{\varphi}^{I_u}(\vec t_\varphi)\wedge P_{\psi}^{I_u}(\vec t_\psi)\indent P_{\varphi\vee\psi}^{I_u}(\vec t_{\varphi\vee\psi}) \Leftrightarrow P_{\varphi}^{I_u}(\vec t_\varphi)\vee P_{\psi}^{I_u}((\vec t_\psi))$$$$ P_{\varphi\to\psi}^{I_u}(\vec t_{\varphi\to\psi}) \Leftrightarrow(\neg P_{\varphi}^{I_w}(\vec t_{\varphi}))\vee P_{\psi}^{I_w}(\vec t_{\psi})\text{ for all $w\geq u$}$$
	$$P_{\forall x\varphi}^{I_u}(\vec t_{\forall x\varphi}) \Leftrightarrow P_{\varphi}^{I_w}(\vec t_\varphi) \text{ for all $w\geq u$, $x\in M_w$}$$$$ P_{\exists x\varphi}^{I_u}(\vec t_{\exists x\varphi}) \Leftrightarrow P_{\varphi}^{I_u}(\vec t_\varphi) \text{ for some $x\in M_u$}$$
	\\
\end{definition}

The above definitions are motivated by the following results which follow directly from the involved definitions and are true both for classical and intuitionistic semantics:

\begin{lemma}
	For every formula $\varphi$ and structure $\mathcal M$
	$$\mathcal S(\mathcal M, \varphi)\models\mathcal S(\varphi)$$
\end{lemma}

\begin{lemma}
	For every structure $\mathcal M$
	$$\mathcal M\models \varphi\text{ iff }S(\mathcal M, \varphi)\models P_\varphi$$
\end{lemma}

Then from the previous three Lemmas we get

\begin{corollary}\label{equivalid}
	In both intuitionistic and classical logic
	\begin{itemize}
		\item $\varphi$ is satisfiable iff $\mathcal \bigwedge S^+(\varphi)\wedge P_\varphi(\vec t_\varphi)$ is. 
		\item $\varphi$ is valid iff $\bigwedge\mathcal S^-(\varphi)\to P_\varphi(\vec t_\varphi)$ is.
	\end{itemize}
\end{corollary}

\section{IPC to CPC}

We are now ready to put everything together. Recall that our goal is for a formula to give a modified formula that is valid classically iff the original formula is valid intuitionistically. As a first step we perform the normalization from the last section, i.e. instead of $\varphi$ we consider $$\varphi' = \bigwedge \mathcal S^-(\varphi)\to P_\varphi$$

Recall that since $\varphi$ is quantifier-free each formula $\psi\in\mathcal S^-(\varphi)$ is of one of the forms
$$\hspace{-0.8cm}A_\psi\to (B_\psi\wedge C_\psi), (A_\psi\wedge B_\psi)\to C_\psi, A_\psi\to (B_\psi\vee C_\psi), (A_\psi\vee B_\psi)\to C_\psi, (A_\psi\to B_\psi)\to C_\psi$$

where we can treat $A\to B$ as a special case $A\to (B\wedge B)$. We then apply the transformation to CQC, leading to a formula
$$\varphi'' = K(\varphi')\to\bigwedge  \mathcal S\to P_\varphi(u)$$
where $u$ is some new free variable and each $\psi\in\mathcal S$ is of one of the forms
$$\hspace*{-2mm}\forall k(u\preceq k\to A(k)\to (B_\psi(k)\wedge C_\psi(k))), \forall k(u\preceq k\to (A_\psi(k)\wedge (B_\psi(k))\to C_\psi(k)))$$
$$\hspace*{-2mm}\forall k(u\preceq k\to A(k)\to (B_\psi(k)\vee C_\psi(k))), \forall k(u\preceq k\to (A_\psi(k)\vee B_\psi(k))\to C_\psi(k))$$
$$\forall k(u\preceq k\to (\forall l(k\preceq l \to A_\psi(l)\to B_\psi(l))\to C_\psi(k)))$$
with each $P(x)$ possibly being $\bot$. Applying Herbrandization we obtain a formula
$$\varphi''' = K(\varphi')\to \bigwedge \mathcal S'\to P_\varphi(s)$$
where $s$ is some new constant term and each $\psi\in\mathcal S'$ is of one of the forms
$$\hspace*{-2mm}\forall k(s\preceq k\to A_\psi(k)\to (B_\psi(k)\wedge C_\psi(k))), \forall k(s\preceq k\to (A_\psi(k)\wedge (B_\psi(k))\to C_\psi(k)))$$
$$\hspace*{-2mm}\forall k(s\preceq k\to A_\psi(k)\to (B_\psi(k)\vee C_\psi(k))), \forall k(s\preceq k\to (A_\psi(k)\vee (B_\psi(k))\to C_\psi(k)))$$
$$\forall k(s\preceq k\to (k\preceq f_\psi(k) \to A_\psi(f_\psi(k))\to B_\psi(f_\psi(k)))\to C_\psi(k))$$
and $f$ is a new function symbol for each introduced formula of that form. Now if $(M, I)$ is a counter-model then we have a counter-model $(M',I')$ where $$M' = \{m\in M|s\preceq m\}$$ and 
$$f_\psi^{I'}(u) = \begin{cases}
	f^I_\psi(u)&\text{ iff $u\preceq f^{I}_\psi(u)$}\\
	u&\text{ else}
\end{cases}$$ that is where $s$ is a least element and the $f_\psi$ increase their argument. Therefore instead of $\varphi'''$ we may consider a formula $$\varphi^\circ = \forall k(s\preceq k)\to \bigwedge\{\forall k(k\preceq f_\psi(k))\:|\:\psi\in\mathcal F_\to\}\to K(\varphi')\to \bigwedge \mathcal S^\circ\to P_\varphi(s)$$ where each formula $\psi\in\mathcal S^\circ$ is of one of the forms
$$\forall k(A_\psi(k)\to (B_\psi(k)\wedge C_\psi(k))), \forall k((A_\psi(k)\wedge B_\psi(k))\to C_\psi(k))$$
$$\forall k(A_\psi(k)\to (B_\psi(k)\vee C_\psi(k))), \forall k((A_\psi(k)\vee B_\psi(k))\to C_\psi(k))$$
$$\forall k((A_\psi(f_\psi(k))\to B_\psi(f_\psi(k)))\to C_\psi(k))$$
and $\mathcal F_\to$ denotes the set of formulas of the last kind.

We now argue that if a counter-model exists for $\varphi^\circ$ then a bounded counter-model exists, allowing us to return to the propositional case.

\begin{definition}
	Suppose we are given a counter-model $\mathcal M = (M, I)$ for $\varphi^\circ$. For $\psi\in\mathcal F_\to$ we say that is it \textit{fulfilled} at $u\in M$ iff $A_\psi^I(u)\to B_\psi^I(u)$ is false or $C_\psi^I(u)$ is true. For $\psi\in\mathcal F_\to$ define $$g_\psi : M\to M, u\mapsto\begin{cases}
		u&\text{ if $\psi$ is fulfilled in $u$}\\
		f^I_\psi(u)&\text{ else}		
	\end{cases}$$Define a $\mathcal M_T = (M_T, I_T)$ as follows:
	\begin{itemize}
		\item $M_T$ is the set of sequences without repetition on $\mathcal F_\to$.
		\item Interpret $\preceq$ as the prefix-order.
		\item Have $$f_\psi^{I_T}(\psi_1\dots\psi_n) = \begin{cases}
			\psi_1\dots\psi_n&\text{if $\psi$ occurs in $\psi_1\dots\psi_n$}\\
			\psi_1\dots\psi_n\psi&\text{else}			
		\end{cases}$$
		\item For propositional variables $P$ have $$P^{I_T}(\psi_1\dots \psi_n) = P^I(g_{\psi_n}(\dots(g_{\psi_1}(s^I))\dots))$$
	\end{itemize}
\end{definition}
\begin{lemma}
	Let $\mathcal M = (M, I)$ be a counter-model to $\mathcal \varphi^\circ$.
	\begin{enumerate}
		\item $\psi$ is fulfilled at $f_\psi^I(u)$ for all $u\in M, \psi\in\mathcal F_\to$.
		\item If $\psi$ is fulfilled at some $u\in M$ then $\psi$ is fulfilled at all $w\geq u$.
	\end{enumerate}

\end{lemma}

\begin{proof}
	1. If $C_\psi^I(u)$ is true then we are done due to persistency. Otherwise we know that $(A_\psi^I(f_\psi^I(u))\to B_\psi^I(f_\psi^I(u)))\to C_\psi^I(f_\psi^I(u))$ holds, so $A_\psi^I(f_\psi^I(u))\to B_\psi^I(f_\psi^I(u))$ must be true.

	2. If $C_\psi^I(u)$ is true then $C_\psi^I(w)$ is also true and we are done. So suppose $A_\psi^I(u)\to B_\psi^I(u)$ is false, i.e. $A_\psi^I(u)$ true and $B_\psi^I(u)$ false. Then due to the persistency condition $A_\psi^I(w)$ is also true. Again if $C_\psi^I(w)$ is true we are done. Otherwise we know that $B^I(f^I_\psi(w))$ must be false, since $(A^I_\psi(f^I_\psi(w))\to B^I(f^I_\psi(w)))\to C^I_\psi(w)$ and $A^I_\psi(f^I_\psi(w))$ hold. But then due to persistency so is $B^I_\psi(w)$ and we have that $\psi$ is fulfilled at $w$.
\end{proof}

Essentially what fulfilment of $\psi$ at $u$ tells us is that having $f_\psi(w) = w$ for all $w\geq u$ works. 
With this lemma in hand showing that $(M_T, I_T)$ is a counter-model to $\varphi^\circ$ is a straightforward check of definitions.

\begin{corollary}
	If $\mathcal M$ is a counter-model to $\varphi^\circ$ then so is $\mathcal M_T$.
\end{corollary}

\begin{example}
	Consider
	$$\mathcal S = \{(A\to B)\to C, (B\to A)\to D, (A\wedge B)\to \bot, (C\vee D)\to E\}$$and $\varphi = \bigwedge S\to E$ which has the Kripke counter-model
	\begin{center}
		\begin{tikzpicture}[squarednode/.style={rectangle, draw, very thick, minimum size=5mm}]
			\node (center) {};
			\node[squarednode]      (right)   [right=of center] {$B, C$};
			\node[squarednode]      (left)    [left=of center] {$A, D$};
			\node[squarednode]      (lower)       	[below=of center] {};
			\draw[-{Stealth[length=3mm, width=3mm]}] (lower.north west) -- (left.south east);
			\draw[-{Stealth[length=3mm, width=3mm]}] (lower.north east) -- (right.south west);			
		\end{tikzpicture}
	\end{center}
	\vspace*{-.3cm}where at each node the true propositions are indicated.\pagebreak

	There is a corresponding classical counter-model $\mathcal M = (M, I)$ to $\varphi^\circ$ with $M = \{u, u_l, u_r\}$, $A^{I}(l) = D^I(l) = B^I(r) = C^I(r) = 1$ and $0$ else.

	Denoting $\psi_1 := (A\to B)\to C$ and $\psi_2 := (A\to B)\to C$ this corresponds to a counter-model $\mathcal M_T = (M_T, I_T)$ of $\varphi^\circ$ with $M_T = \{\epsilon, \psi_1, \psi_2, \psi_1\psi_2, \psi_2\psi_1\}$ and interpretation as presented below
	\begin{center}
		\begin{tikzpicture}[squarednode/.style={rectangle, draw, very thick, minimum size=5mm}]
			\node (center) {};
			\node[squarednode]      (right1)   [right=of center] {$B, C$};
			\node[squarednode]      (right2)   [above=of right1] {$B, C$};
			\node[squarednode]      (left1)    [left=of center] {$A, D$};
			\node[squarednode]      (left2)    [above=of left1] {$A, D$};
			\node[squarednode]      (lower)       	[below=of center] {};
			\draw[-{Stealth[length=3mm, width=3mm]}] (lower.north west) -- (left1.south east) node [midway, below left] {$f_{\psi_1}$};
			\draw[-{Stealth[length=3mm, width=3mm]}] (lower.north east) -- (right1.south west) node [midway, below right] {$f_{\psi_2}$};
			\draw[-{Stealth[length=3mm, width=3mm]}] (left1.north) -- (left2.south) node [midway, right] {$f_{\psi_2}$};
			\draw[-{Stealth[length=3mm, width=3mm]}] (right1.north) -- (right2.south) node [midway, left] {$f_{\psi_1}$};
			\draw[-{Stealth[length=3mm, width=3mm]}] (left1.south west) to [out=225, in=135, loop, looseness=3] node [left] {$f_{\psi_1}$} (left1.north west);		
			\draw[-{Stealth[length=3mm, width=3mm]}] (right1.south east) to [out=315, in=45, loop, looseness=3] node [right] {$f_{\psi_2}$} (right1.north east);
			\draw[-{Stealth[length=3mm, width=3mm]}] (left2.north west) to [out=135, in=45, loop, looseness=3] node [midway, above] {$f_{\psi_1}, f_{\psi_2}$} (left2.north east);
			\draw[-{Stealth[length=3mm, width=3mm]}] (right2.north west) to [out=135, in=45, loop, looseness=3] node [midway, above] {$f_{\psi_1}, f_{\psi_2}$} (right2.north east);		
		\end{tikzpicture}
	\end{center}
\end{example}

So we know that $\varphi^\circ$ is valid iff it is valid for structures that have as a domain the sequences without repetition over $\mathcal F_\to$, ordered by the prefix-relation and $$s^I = \epsilon, f_\psi^I(x) = \begin{cases}
	x&\text{ if $\psi$ occurs in $x$}\\
	x\psi&\text{ otherwise}
\end{cases}$$ Now we can replace all $\forall$-quantifiers by enumerating all the ground terms and finally all distinct ground instances of relations by new propositional variables yielding a CPC formula. This gives us the following translation:

\begin{definition}
	Suppose $\mathcal S$ is a set of formulas of the form
	$$A\to (B\wedge C), (A\wedge B)\to C, A\to (B\vee C),(A\vee B)\to C, (A\to B)\to C$$
	Let $\mathcal F_\to$ be the set of formulas of the form $(A\to B)\to C$ and $\Lambda$ the set of sequences without repetition over $\mathcal F_\to$. For each atom $A$ occurring in $\mathcal S$ and $k\in \mathcal F_\to$ consider a new atom $A^{k}$, which will correspond to $A^{I_F}(k)$ from the previous definition. Obtain $\mathcal S^\#$ by including the formulas
	\begin{itemize}
		\item $A^k\to A^{k\psi}$ for $A$ occurring in $\mathcal S$, $k\in\Lambda$ and $\psi\in\mathcal F_\to$ not occurring in $k$.
		\item $A^k\to (B^k\circ C^k)$ for each $\circ\in\{\wedge,\vee\}$, $A\to (B\circ C)\in\mathcal S$, $k\in\Lambda$.
		\item $(A^k\circ B^k)\to C^k$ for each $\circ\in\{\wedge,\vee\}$, $A\to (B\circ C)\in\mathcal S$, $k\in\Lambda$.
		\item $(A^{k\psi}\to B^{k\psi})\to C^k$ for $\psi = (A\to B)\to C\in\mathcal S$, $k\in\Lambda$ if $\psi$ does not occurr in $k$.
	\end{itemize}
\end{definition}

\begin{theorem}
	$$\mathcal S\models_C P\text{ iff }\mathcal S^\#\models_I P^\emptyset$$
\end{theorem}

\begin{example}
	Let us consider the law of excluded middle $\varphi = A\vee\neg A$. Then we get
	$$\mathcal S^-(\varphi) = \{(P_A\vee P_{\neg A})\to P_\varphi, A\to P_A, (A\to \bot)\to P_{\neg A}\}$$
	and we know that $\varphi$ is valid iff $\bigwedge \mathcal S(\varphi)\to P_\varphi$ is valid. Applying the above lemma this is intuitionistically valid iff $\bigwedge\mathcal S^\#\to P^\emptyset_\varphi$ is classically where, setting $\psi = (A\to \bot)\to P_{\neg A}$,
	\begin{align*}
		\mathcal S^\# =&\{P_\varphi^\epsilon\to P_\varphi^{\psi}, P_A^\epsilon\to P_A^{\psi},P_{\neg A}^\epsilon\to P_{\neg A}^{\psi},A^\epsilon\to A^{\psi}\}\\ &\cup\{(P_A^\epsilon\vee P_{\neg A}^\epsilon)\to P_\varphi^\epsilon, (P_A^{\psi}\vee P_{\neg A}^{\psi})\to P_\varphi^{\psi},  A^\epsilon\to P_A^\epsilon\}\\ &\cup \{P_A^{\psi}\to P_A^{\psi}, (A^{\psi}\to \bot)\to P_{\neg A}^\epsilon, (A^{\psi}\to \bot)\to P_{\neg A}^{\psi}\}
	\end{align*}
	It is easy to construct a counter example:
	$$P_\varphi^\epsilon  = P_A^\epsilon = P_{\neg A}^\epsilon = A^\epsilon = 0$$
	$$P_{\varphi}^{\psi} = P_A^{\psi} =  P_{\neg A}^{\psi} = A^{\psi} = 1$$
\end{example}

\section{IPC to QBF}

It is well known that a polynomial time translation between QBF and IPC must exists, since both problems are PSPACE complete~\cite{garey1979computers, statman1979intuitionistic}. However giving an explicit translation from IPC to QBF can still be useful as good solvers exists for QBF but for IPC not so much. Furthermore we will be able to give fragments of IPC for which validity is complete for other stages of the polynomial hierarchy. The translation from the last section will serve as a useful tool for this. We will not explain the semantics of QBF in depth. Intuitively just think that the interpretation of existentially quantified variables may depend on variables quantified above them.

Fix some set $\mathcal S$ of formulas $\psi$ of the form $$\hspace*{-.7cm}A_\psi\to (B_\psi\wedge C_\psi), (A_\psi\wedge B_\psi)\to C_\psi, A_\psi\to (B_\psi\vee C_\psi), (A_\psi\vee B_\psi)\to C_\psi, (A_\psi\to B_\psi)\to C_\psi$$and a formula $$\varphi = \bigwedge \mathcal S\to P$$
and denote with $\mathcal F_\to$ the set of formulas of the form $(A\to B)\to C$. We will now formulate a QBF that expresses that $\varphi$ has an intuitionistic counter-model. The idea how to get a linear-size formula is: Instead of expressing validity at every node of $\mathcal M_T$ - which was done in the CPC translation - we express that on each path in $\mathcal M_T$ the nodes satisfy all conditions for $\mathcal M_T$ being a counter-model. The universally quantified variables $F_\psi$ in the next definition handle the branching, i.e. express which path is considered.

\begin{definition}
	For every propositional variable $X$ occurring in $\mathcal S$ and non-negative integer $n$ consider a new atom $X^n$ and for every formula $\psi\in\mathcal F_\to$ consider a new atom $F_\psi^n$. Let $\vec X^n$ range over all propositional variables $X^n$ and $\vec F^n$ range over all $F_\psi^n$. Define
	\begin{itemize}
		\item valid$(n)$ which encodes that $\vec F^n_\psi$ represents a valid next element of a sequence without repetition, i.e. if we interpret the $\vec F^i$ as bit-vectors, then $\vec F^n$ has exactly one bit set to $1$, indicating the formula at the $n$-th position of the sequence, and it is at a position that is $0$ in $\sum\{\vec F^i\:|\:i < n\}$.
		\item per($n$) which encodes that the persistency condition holds.
		\item sat$_{\mathcal S}(n)$ encoding that the formulas in $\mathcal S\setminus\mathcal F_\to$ hold at the $n$-th world of the path.
		\item sat$_{\mathcal F_\to}(n)$ encoding that the formula in $\mathcal F_\to$ that is represented by $\vec F^{n-1}$ holds.
	\end{itemize}
	Have $k = |\mathcal F_\to|$ and define
	$$\varphi^Q_{k} = \exists \vec X^k\left(\text{per}(k)\wedge \text{sat}_{\mathcal S}(k)\wedge \text{sat}_{\mathcal F_\to}(k)\right)$$
	for $0 < i < k$
	$$\hspace*{-.8cm}\varphi^Q_{i} = \exists \vec X^i\forall \vec F^i\left(\text{per}(i)\wedge \text{sat}_{\mathcal S}(i)\wedge \text{sat}_{\mathcal F_\to}(i)\wedge\left(\text{valid}(i)\to \varphi^Q_{i+1}\right)\right)$$
	and $$\varphi^Q = \varphi_0^Q = \exists \vec X^0\forall \vec F^0\left(\neg P^0\wedge \text{sat}_{\mathcal S}(0)\wedge(\text{valid}(0)\to \varphi_1^Q)\right)$$
	Example encodings of the above formulas would be
	\begin{align*}
		\text{valid}(n) = &\left(\bigvee \{\mathcal F^n_\psi\:|\:\psi\in\mathcal F_\to\}\right)\wedge\left(\bigwedge \{\neg(F^n_{\psi_1}\wedge F^n_{\psi_1})\:|\:\psi_1\neq\psi_2\in\mathcal F_\to\}\right)\wedge\\&\wedge\left(\bigwedge\left\{\bigwedge\{F^i_\psi\to\neg F^n_\psi\:|\:i < n\}\:|\:\psi\in\mathcal F_\to\right\}\right)\\
		\text{per}(n) = & \bigwedge\{X^{n-1}\to X^n\:|\: X \text{ prop. variable with }X^{n-1}\in\vec X^{n-1}, X^n \in\vec X^{n}\}\\
		\text{sat}_{\mathcal S}(n) = &\bigwedge\left\{(A^n\circ B^n)\to C^n\:|\:(A\circ B)\to C\in \mathcal S\setminus\mathcal F_\to\right\}\wedge\\
		&\bigwedge\left\{A^n\to (B^n\circ C^n)\:|\:A\to (B\to C)\in \mathcal S\setminus\mathcal F_\to\right\}\\
		\text{sat}_{\mathcal F_\to}(n) = & \bigwedge\{F^{n-1}_\psi\to (A^n\to B^n)\to C^{n-1}\:|\:\psi = (A\to B)\to C\in\mathcal F_\to\}
	\end{align*}
\end{definition}

\begin{example}
	With the above encodings for double negation elimination $\varphi = ((A\to \bot)\to \bot)\to A$ we have
	$$\hspace*{-.3cm}
	\varphi^Q = \exists A^0\forall F^0\left(\neg A^0\wedge\left(F^0\to \exists A^1\left((A^0\to A^1)\wedge(F_0\to (A^1\to \bot)\to \bot)\right)\right)\right)
	$$
	which is a satisfiable QBF since $A^0 = 0, A^1 = 1$ satisfies it for any $F^0$.
\end{example}

\begin{lemma}
	$\varphi$ is not intuitionistically valid iff $\varphi^Q$ is a satisfiable QBF.
\end{lemma}
\begin{proof}
	Suppose $\varphi$ is not intuitionistically valid, i.e. there exists a counter-model $\mathcal M$ for $\mathcal S^\#\to P^\epsilon$. For each atom $A$ interpret $A^0$ such as $\mathcal M$ interprets $A^\epsilon$. Suppose we are given interpretations of all atoms $A^i$ for $i < n$ and a sequence with no repetitions $\psi_1\dots\psi_{n-1}$ over $\mathcal F_\to$ such that $\psi_i$ is exactly the $\psi\in\mathcal F_\to$ for which $F_{\psi}^i$ is true and $A^i$ is interpreted as $A^{\psi_1\dots\psi_i}$ is in $\mathcal M$.

	Let the $F^{n}_\psi$ be arbitrarily interpreted (since they are $\forall$-quantified). If not exactly one is interpreted as true then valid$(n)$ fails, i.e. have $\psi_n\in\mathcal F_\to$ the only element with $F^n_{\psi_n} = 1$. If $F^i_{\psi_n} = 1$ for some $i < n$ then valid$(n)$ also fails, so we may assume that $\psi_1\dots\psi_n$ is a sequence with no repetitions. Interpret the atoms $A^n$ as $\mathcal M$ interprets $A^{\psi_1\dots\psi_n}$. Continue this construction until $n  = |\mathcal F_\to|$. Then from $\mathcal M$ being a counter-example to $\mathcal S^\#\to P$ it directly follows that this interpretation satisfies $\varphi^Q$.

	On the other hand suppose $\varphi^Q$ is satisfiable. We construct a counter-example to $\varphi^\#$.
	Again we proceed iteratively. Interpret $A^\epsilon$ such as $A^0$ is interpreted in some satisfying interpretation of $\varphi^Q$. Suppose we are given a sequence $\psi_1\dots \psi_{n-1}$ such that for $i<n$ having $A^i = A^{\psi_0\dots\psi_{n-1}}$ is part of a satisfying interpretation of $\varphi^Q$ in which $F^i_\psi$ is chosen true iff $\psi = \psi_i$. Let $\psi_n\in\mathcal F_\to$. Consider some interpretation of the $A^n$ that are part of a satisfying assignment where $F^n_\psi$ is true iff $\psi = \psi_n$ and all variables quantified above are chosen as before. Have $A^{\psi_0\dots\psi_n} = A^n$ for each propositional variable $A$. From the definitions it directly follows that construction yields a counter-model for $\varphi^\#$.
\end{proof}

Since we have not discussed QBF in detail the above argument might be a bit difficult to follow, in particular the second part, but is simple in essence. We shall demonstrate with help of a example.

\begin{example}
	Consider the formula $\varphi = (((A\to B)\to C)\wedge((B\to A)\to C))\to C$, which is a non-intuitionistic tautology. Its associated QBF is $$\varphi^Q = \exists A^0B^0C^0\forall F_{\psi_1}^1F_{\psi_2}^1 \exists A^1B^1C^1\forall F_{\psi_1}^2F_{\psi_2}^2 \exists A^2B^2C^2\psi'$$where $\psi_1 := (A\to B)\to C, \psi_2 = (B\to A)\to C$ and $$\psi' = \bigwedge\left\{\begin{matrix}
		F^1_{\psi_1}\vee F^1_{\psi_2}, \neg(F^1_{\psi_1}\wedge F^1_{\psi_2}), F^2_{\psi_1}\vee F^2_{\psi_2},\neg(F^2_{\psi_1}\wedge F^2_{\psi_2}),\\F^1_{\psi_1}\to\neg F^2_{\psi_1},F^1_{\psi_2}\to\neg F^2_{\psi_2}, A^0\to A^1,A^1\to A^2, 
		\\B^0\to B^1, B^1\to B^2, C^0\to C^1, C^1\to C^2,\neg C^0, \\
		F^1_{\psi_1}\to (A^1\to B^1)\to C^0, F^1_{\psi_2}\to(B^1\to A^1)\to C^0\\
		F^2_{\psi_1}\to (A^2\to B^2)\to C^1, F^2_{\psi_2}\to(B^2\to A^2)\to C^1
	\end{matrix}\right\}$$
	This QBF is satisfiable: Have $P^0 = 0$ for all atoms $P$, always $P^2 = 1$ for all atoms $P$ and $C^1 = 1$. Furthermore have $A^1 = 1$ iff $F^1_{\psi_1} = 1$ and $B^1 = 1$ iff $F^1_{\psi_2} = 1$. The construction of an counter-model to $\varphi^\#$ then proceeds as follows:
	\begin{itemize}
		\item First have $P^\epsilon = 0$ for each propositional variable $P$.
		\item For $P^{\psi_1}$ we consider what happens when we choose $F^1_{\psi_1}$ to be true, i.e. we have $A^{\psi_1} = C^{\psi_1} = 1$ and $0$ for all other propositional variables.
		\item For $P^{\psi_2}$ we consider what happens when we choose $F^1_{\psi_2}$ to be true, i.e. we have $B^{\psi_1} = C^{\psi_1} = 1$ and $0$ for all other propositional variables.
		\item Finally for $P^{\psi_1\psi_2} = P^{\psi_2\psi_1} = 1$ for all propositional variables $P$ as $P^2$ is true regardless of choice for the $F^i_\psi$. 
	\end{itemize}
\end{example}

Note that the last set of existentially quantified variables is superfluous because there is only one assignment which even has the chance to falsify $\varphi^Q$, namely that which assigns $1$ to the only $F_\psi^n$ such that $F_\psi^i = 0$ for all $i < n$. That is in our final formula we can replace $F_\psi^n$ with $\bigwedge\{\neg F_\psi^i\:|\:i < n\}$ and remove that quantification over $F_\psi^i$. With that we get the following:

\begin{corollary}
	Let $N$-INTINVALID be the problem of deciding if a formula $\bigwedge \mathcal S\to P$ as above with $|\mathcal F_\to|\leq N$ is not intuitionistically valid. Then $N$-INTINVALID is in $\Sigma_{2N-1}^P$. The dual problem $N$-INTVALID is in $\Pi_{2N-1}^P$.
\end{corollary}

As a final result we would like to prove that the above problems are complete for their respective complexity classes. It is sufficient to provide a polynomial reduction from QBF with bounded number of quantifier alternation to them. Such translations exists~\cite{statman1979intuitionistic,vsvejdar2003polynomial}, however they are insufficient in that they generate more formulas of form $(A\to B)\to C$ than there are quantifier alternations. We therefore would need a slightly new translation.

\section{CQC to IQC}

We now give an analogous transformation for first order logic. However our trick of lifting the sentence to first-order logic to encode the Kripke semantics no longer works in such a straightforward manner, since we already are in first-order logic. However we do it nonetheless! The domain of our classical model will feature on the one hand elements representing worlds in a Kripke frame and on the other hand the domain-elements from the Kripke model. We reconcile these notions by introducing a special binary predicate $E$, first considered in~\cite{baaz2006skolemization} and expanded onto in~\cite{iemhoff2010eskolemization}, where $E(x, u)$ encodes that $x$ is an element of the domain of the world $u$. In addition we use a unary predicate $W$ that encodes that an element represents a world.

First of all recall that due to~\ref{equivalid} it is sufficient to consider sentences of the form $\bigwedge\mathcal S\to P$ where all $\psi\in\mathcal S$ are of one of the forms

$$\forall \vec t(A(\vec t_{A})\to (B(\vec t_B)\wedge C(\vec t_C))), \forall \vec t((A(\vec t_{A})\wedge B(\vec t_B))\to C(\vec t_C)),$$$$ \forall \vec t(A(\vec t_{A})\to (B(\vec t_B)\vee C(\vec t_C))),
\forall \vec t((A(\vec t_{A})\vee B(\vec t_B))\to C(\vec t_C)),$$$$ \forall \vec t((A(\vec t_A)\to B(\vec t_B))\to C(\vec t_C)),\forall \vec t(\forall xA(\vec t_A)\to B(\vec t_B)),$$$$ \forall \vec t(A(\vec t_A)\to\forall xB(t_B)), \forall \vec t(\exists xA(\vec t_A)\to B(\vec t_B)), \forall \vec t(A(\vec t_A)\to\exists xB(t_B))$$	
First let us note how we axiomatize $E$ and $W$. For a vector $\vec x = x_1\dots x_n$ we define $\vec E(\vec x, u) = \bigwedge\{E(x_i, u)\:|\:i\leq n\}$.
\begin{align*}
	E(\varphi) = & \forall u\forall x(E(x, u)\to(\neg W(x)\wedge W(u)))\wedge\\
	& \forall u\forall w\forall x((E(x, u)\wedge u\preceq w)\to E(x, w))\wedge\\
	& \bigwedge\{\forall\vec t\forall u(\vec E(\vec t, u)\to E(f(\vec t), u))\:|\:\text{$f$ is an $n$-ary function symbol}\}
\end{align*}
The one possible problem in this encoding is that we have a global interpretation of function symbols, i.e. in an intuitionistic model we could have worlds $u, v$ such that $f^{I_u}\neq f^{I_v}$. However this is not a problem: We can take as the domain of our classical model equivalence classes $\{[(x, u)]\:|\: x\in M_u\}$ under the relation $(x, u)\sim (y, v)$ iff $x = y$ and there exists $w$ comparable with $u, v$ such that $x\in w$ and interpret $f^I([x, u]) = [f^{I_u}(x), u]$. Then our classical translation only forces functions to be equal on arguments on which they must be equal anyway. So the Kripke model does in fact not afford us any additional generality.

As before we need to axiomatize $\preceq$. For each $n$-ary relation symbol consider $P$ a new $n+1$ ary relation symbol $P^\#$.  The only difference in that for persistence we may require additional quantification. 
\begin{align*}
	K(\varphi) = & \forall k(k\preceq k)\wedge \forall k\forall l((k\preceq l\wedge l\preceq k)\to l = k)\wedge\\
	& \forall k\forall l\forall m((k\preceq l\wedge l\preceq m)\to k\preceq m)\wedge\\
	& \bigwedge\{\forall k\forall l(k\preceq l\to \forall \vec x(A^\#(\vec x, k)\to A^\#(\vec x, l)))\:|\:\text{$A$ occurs in $\varphi$}\}
\end{align*}
Using $E$ and $W$ we obtain a set $\mathcal S^\circ$ by including formulas as follows:

\begin{itemize}
	\item $\forall \vec t\forall u(s\preceq u\to\vec E(\vec t, u)\to (A_\psi^\#(t_A, u)\to (B_\psi^\#(\vec t_B, u)\circ C_\psi^\#(\vec t_C, u))))$\\for each $\circ\in\{\wedge, \vee\}, \psi = \forall \vec t(A_\psi(\vec t_A)\to (B_\psi(\vec t_B)\circ C_\psi(\vec t_C)))\in\mathcal S$
	\item $\forall \vec t\forall u(s\preceq u\to\vec E(\vec t, u)\to \forall w(u\preceq w\to A_\psi^\#(\vec t_A, w)\to B_\psi^\#(\vec t_B, w))\to C_\psi^\#(\vec t_C, u))$ for each $\psi = \forall \vec t((A_\psi(\vec t_A)\to B_\psi(\vec t_B))\to C_\psi(\vec t_C))\in\mathcal S$
	\item  $\forall \vec t\forall u(s\preceq u\to\vec E(\vec t, u)\to \forall w(u\preceq w\to \forall x(E(x, w)\to A_\psi^\#(\vec t_A, w)))\to B^\#(\vec t_B, u))$ for each $\forall \vec t(\forall xA_\psi(\vec t_A)\to B_\psi(\vec t_B))\in\mathcal S$.
	\item $\forall \vec t\forall u(s\preceq u\to\vec E(\vec t, u)\to A^\#_\psi(\vec t_A, u)\to \forall x(E(x, u)\to B^\#_\psi(\vec t_B, u)))$\\for each $\forall \vec t(A(\vec t_A)\to \forall xB(\vec t_B))\in\mathcal S$.
	\item $\forall \vec t\forall u(s\preceq u\to\vec E(\vec t, u)\to \exists x(E(x, u)\wedge A^\#_\psi(\vec t_A, u))\to B^\#_\psi(\vec t_B))$\\for each $\forall \vec t(\exists xA_\psi(\vec t_A)\to B_\psi(\vec t_B))\in\mathcal S$
	\item $\forall \vec t\forall u(s\preceq u\to\vec E(\vec t, u)\to A^\#_\psi(\vec t_B)\to \exists x(E(x, u)\wedge B^\#_\psi(\vec t_B, u)))$\\for each $\forall \vec t(A_\psi(\vec t_A)\to \exists xB_\psi(\vec t_B))\in\mathcal S$
\end{itemize}
Then $\varphi = \bigwedge\mathcal S\to P$ is intuitionistically valid iff
$$\varphi^\circ= E(\varphi)\to K(\varphi)\to \bigwedge S^\circ\to P^\#(s)$$
is classically valid.

We proceed now as in the previous section. Denote the set of formulas $\psi\in\mathcal S$ of the form $\forall\vec t((A_\psi(\vec t_A)\to B_\psi(\vec t_B))\to C_\psi(\vec t_C))$ with $\mathcal F_\to$ and of the form $\forall \vec t(\forall xA_\psi(\vec t_A)\to B_\psi(\vec t_B))$ with $\mathcal F_\forall$. Let $\mathcal F = \mathcal F_\to\cup\mathcal F_\forall$. For each  $\psi\in\mathcal F$ consider a new function symbol $f_\psi$. As before we may assume that $s$ is a least element and eliminate the quantification over $w$ by replacing formulas
$$\forall \vec t\forall u(s\preceq u\to\vec E(\vec t, u)\to \forall w(u\preceq w\to A_\psi^\#(\vec t_A, w)\to B_\psi^\#(\vec t_B, w))\to C_\psi^\#(\vec t_C, u))$$ with
$$\forall \vec t\forall u(\vec E(\vec t, u)\to (A_\psi^\#(\vec t_A, f_\psi(\vec t, u))\to B_\psi^\#(\vec t_B, f_\psi(\vec t, u)))\to C_\psi^\#(\vec t_C, u))$$ and
$$\forall \vec t\forall u(s\preceq u\to\vec E(\vec t, u)\to \forall w(u\preceq w\to \forall x(E(x, w)\to A_\psi^\#(\vec t_A, w)))\to B^\#(\vec t_B, u))$$ with $$\forall \vec t\forall u(\vec E(\vec t, u)\to (\forall x(E(x, f_\psi(\vec t, u))\to A_\psi^\#(\vec t_A, f_\psi(\vec t, u))))\to B^\#(\vec t_B, u))$$
and in each other formula in $\mathcal S^\circ$ remove the $s\preceq u$ to obtain $\mathcal S^\#$ from $\mathcal S$. Then the previous formulas are equivalid to
$$\hspace*{-.5cm}\varphi^\circ= \forall u(s\preceq u)\to \bigwedge\{\forall \vec t\forall u(u\preceq f_\psi(\vec t, u))\:|\:\psi\in\mathcal F\}\to E(\varphi)\to K(\varphi)\to\bigwedge\mathcal S^\#\to P^\#$$
As in the previous section we can define a tree-counter model for this.
\begin{definition}
	Suppose we are given a counter-model $\mathcal M = (M, I)$ for $\varphi^\circ$. For $\psi = \forall\vec t\psi'\in\mathcal F_\to$ we say that is it \textit{fulfilled} at $u\in M$ with $\vec a\in M^n$ iff $(A_\psi^I(\vec t_A, u)\to B_\psi^I(\vec t_B, u))[\vec a/\vec t]$ is false or $C_\psi^I(\vec t_C, u)[\vec a/\vec t]$ is true.
	For $\psi = \forall\vec t\psi'\in\mathcal F_\forall$ we say that it is \textit{fulfilled} at $u\in M$ with $\vec a\in M^n$ iff there is $a\in M_u$ such that $A^{I[a/x]}_\psi(\vec t_A)[\vec a/\vec t]$ is false or $B_\psi^I(\vec t_B)[\vec a/\vec t]$ is true. For $\psi\in\mathcal F$ define $$g_{\vec a, \psi} : M\to M, u\mapsto\begin{cases}
		u&\text{ if $\psi$ is fulfilled in $u$ with $\vec a$}\\
		f^I_\psi(\vec a, u)&\text{ else}		
	\end{cases}$$		
	Define a $\mathcal M_T = (M_T, I_T)$ as follows:
	\begin{itemize}
		\item $M_T$ are the sequences on $\{ \vec t, \psi\:|\:\psi\in \mathcal F, \vec t\in M^n\}$ without repetition of elements $\vec t,\psi$ with $\psi\in\mathcal F_\to$ and without immediate repetition of elements $\vec t,\psi$ with $\psi\in\mathcal F_\forall$.
		\item Interpret $\preceq$ as the prefix-order.
		\item Have $$f_\psi^{I_T}(x_1\dots x_n) = \begin{cases}
			x_1\dots x_n&\text{if $x=x_n$}\\
			x_1\dots x_nx&\text{else}			
		\end{cases}$$
		\item For propositional variables $P$ have $$P^{I_T}(x_1\dots x_n) = P^I(g_{x_n}(\dots(g_{x_1}(s^I))\dots))$$
	\end{itemize}
\end{definition}

\begin{lemma}
	Let $\mathcal M = (M, I)$ be a counter-model to $\mathcal \varphi^\circ$.
	\begin{enumerate}
		\item $\psi$ is fulfilled at $f_\psi^I(\vec a, u)$ with $\vec a$ for all $u\in M, \psi\in\mathcal F$, $\vec a\in M^n$.
		\item If $\psi\in\mathcal F_\to$ is fulfilled at $u\in M$ with $\vec a$ then $\psi$ is fulfilled at all $w\geq u$ with $\vec a$.
	\end{enumerate}	
\end{lemma}

\begin{proof}
	For $\psi\in\mathcal F_\to$ the proof is analogous to the previous section.

	1. Let $\psi\in\mathcal F_\forall$. If $B^I_\psi(\vec t_B, u)[\vec a/\vec t]$ holds then we are done due to persistency. Otherwise there is $a$ with $E^I(a, f^I_\psi(\vec a, u))$ such that $A^{I[a/x]}_\psi(\vec t_A, f^I_\psi(\vec a, u))$ is false. So $\psi$ is fulfilled at $f^I_\psi(\vec a, u)$ with $\vec a$.
\end{proof}

That the second point does not work for $\psi\in\mathcal F_\forall$ is why we must switch to immediate repetition. It does not work because even if for all $x$ with $E(x, w$) we have $A(x, w)$  there could be some term $t$ with $\neg E(t, w)$ and $v$ with $w\preceq v$ such that $E(t, v)$ and $\neg A(t, w)$.

\begin{example}
	Consider the $\varphi = \neg(\neg \forall xA(x)\vee\neg \forall xB(x))$. There is a counter-model
	\begin{center}
		\begin{tikzpicture}[squarednode/.style={rectangle, draw, very thick, minimum size=5mm}]
			\node[squarednode]      (0)   [] {$ $};
			\node[squarednode]      (1r)   [above right=of 0] {{\color{blue}$1$}};
			\node[squarednode]      (1l)   [above left=of 0] {{\color{red}$1$}};
			\node[squarednode]      (2r)   [above=of 1r] {{1, \color{blue}$2$}};
			\node[squarednode]      (2l)   [above=of 1l] {{1, \color{red}$2$}};
			\node[squarednode]      (3r)   [above=of 2r] {{1, 2, \color{blue}$3$}};
			\node[squarednode]      (3l)   [above=of 2l] {{1, 2, \color{red}$3$}};
			\node      (4r)   [above=of 3r] {{$\vdots$}};
			\node      (4l)   [above=of 3l] {{$\vdots$}};
			\draw[-{Stealth[length=3mm, width=3mm]}, color=blue] (0.north east) -- (1r.south west);
			\draw[-{Stealth[length=3mm, width=3mm]}, color=red] (0.north west) -- (1l.south east);
			\draw[-{Stealth[length=3mm, width=3mm]}, color=blue] (1r.north) -- (2r.south);
			\draw[-{Stealth[length=3mm, width=3mm]}, color=red] (1l.north) -- (2l.south);
			\draw[-{Stealth[length=3mm, width=3mm]}, color=blue] (1l.north east) -- (2r.south west);
			\draw[-{Stealth[length=3mm, width=3mm]}, color=red] (1r.north west) -- (2l.south east);
			\draw[-{Stealth[length=3mm, width=3mm]}, color=blue] (2r.north) -- (3r.south);
			\draw[-{Stealth[length=3mm, width=3mm]}, color=red] (2l.north) -- (3l.south);
			\draw[-{Stealth[length=3mm, width=3mm]}, color=blue] (2l.north east) -- (3r.south west);
			\draw[-{Stealth[length=3mm, width=3mm]}, color=red] (2r.north west) -- (3l.south east);
			\draw[-{Stealth[length=3mm, width=3mm]}, color=blue] (3r.north) -- (4r.south);
			\draw[-{Stealth[length=3mm, width=3mm]}, color=red] (3l.north) -- (4l.south);
			\draw[-{Stealth[length=3mm, width=3mm]}, color=blue] (3l.north east) -- (4r.south west);
			\draw[-{Stealth[length=3mm, width=3mm]}, color=red] (3r.north west) -- (4l.south east);
		\end{tikzpicture}
	\end{center}
	where at each node the elements that exist at it are indicated. For black ones both $A$ and $B$ hold, for red ones only $B$, for blue ones only $A$. Furthermore red arrows indicate $f_{\neg \forall xA(x)}$ and blue arrows $f_{\neg \forall xB(x)}$. Note that applying $f_{\neg \forall xA(x)}$ we must always reach the left column and $f_{\neg \forall xB(x)}$ we must always reach the right column. Therefore applying $f_{\neg \forall xA(x)}\circ f_{\neg \forall xB(x)}$ we can never remain stationary. This is in contrast to the propositional case.
\end{example}

\begin{corollary}
	If $\mathcal M$ is a counter-model to $\varphi^\circ$ then so is $\mathcal M_T$.
\end{corollary}

Even if we don't have the nice finiteness of the propositional case we have still reduced the model complexity in that we know that considering worlds of the form $f_{\psi_1}(\vec t_1, f_{\psi_2}(\vec t_2, \dots f_{\psi_n}(\vec t_n, s)\dots))$ is sufficient. In particular this allows us to completely eliminate $\preceq$ from the formula, that is we replace $\forall u\forall w\forall x((E(x, u)\wedge u\preceq w)\to E(x, w))$ with $\bigwedge\{\forall \vec t\forall u(E(x, u)\to E(x, f_\psi(\vec t, u)))\:|\:\psi\in\mathcal F_\to\}$ in $E(\varphi)$ to get $E^\#(\varphi)$ and obtain a formula

$$\varphi^{\#} = E^\#(\varphi)\to \bigwedge S^\#\to P^\#$$

\begin{theorem}
	$\varphi$ is intuitionistically valid iff $\varphi^\#$ is classically valid.
\end{theorem}

\begin{example}
	Consider the double negation shift (DNS)
	$$\varphi = \forall x\neg\neg A(x)\to \neg\neg\forall x A(x)$$
	which is a famous non-intuitionistic tautology. We get
	\begin{align*}
		\mathcal S = & \{(P_{\forall x\neg\neg A(x)}\to P_{\neg\neg\forall xA(x)})\to P_\varphi, P_{\forall x\neg\neg A(x)}\to \forall xP_{\neg\neg A(x)}(x)\}\cup\\
		& \{\forall t((P_{\neg\neg A(x)}(t)\wedge P_{\neg A(x)}(t))\to \bot), \forall t((A(t)\to \bot)\to P_{\neg A(x)})(t)\}\cup\\
		& \{(P_{\neg\forall xA(x)}\to \bot)\to P_{\neg\neg\forall xA(x)}, (P_{\neg\forall xA(x)}\wedge P_{\forall xA(x)})\to \bot\}\cup\\
		&\{\forall xA(x)\to P_{\forall xA(x)}\}
	\end{align*}
	and thus, denoting $\psi_1 = (P_{\neg\forall xA(x)}\to \bot)\to P_{\neg\neg\forall xA(x)}$ and $\psi_2 = \forall xA(x)\to P_{\forall xA(x)}$, we have
	\begin{align*}
		\mathcal S^\# = & \{\forall u(P_{\forall x\neg\neg A(x)}(u)\to P_{\neg\neg\forall xA(x)}(u))\to P_\varphi(u)\}\cup\\& \{\forall u(P_{\forall x\neg\neg A(x)}(u)\to \forall xP_{\neg\neg A(x)}(x, u)\}\cup\\
		& \{\forall t, u((P_{\neg\neg A(x)}(t, u)\wedge P_{\neg A(x)}(t, u))\to \bot)\}\\& \{\forall t, u((A(t, u)\to \bot)\to P_{\neg A(x)}(t, u))\}\cup\\
		& \{\forall u((P_{\neg\forall xA(x)}(f_{\psi_1}(u))\to \bot)\to P_{\neg\neg\forall xA(x)}(u))\}\\& \{\forall u((P_{\neg\forall xA(x)}(u)\wedge P_{\forall xA(x)}(u))\to \bot)\}\cup\\
		&\{\forall u(\forall xA(x, f_{\psi_2}(u))\to P_{\forall xA(x)}(u))\}
	\end{align*}
\end{example}

\bibliography{lipics-v2021-EIICL}

\appendix

\begin{proof}[Lemma \ref{ap1}]
	We proceed by simultaneous induction on the formula height.

	For atomic $\chi$ the claims are clear.

	Otherwise there are $3$ cases:

	1. $\chi = \varphi\circ\psi$ for $\circ\in\{\wedge,\vee,\to\}$. By induction hypothesis there exist $\mathcal M_{\varphi, v} = (M, I_{\varphi, v}), \mathcal M_{\psi, v} = (M, I_{\psi, v})$ for $\varphi,\psi$ as in the lemma. Then we can define $I_{\chi, v}$ as follows:

	For every symbol $p$ that occurs in $\varphi^S$ or $\varphi^H$ but not $\psi^S, \psi^H$ have $p^{I_{\chi, v}} = p^{I_{\varphi, v}}$. For every symbol $p$ that occurs in $\psi^S$ or $\psi^H$ but not $\varphi^S, \varphi^H$ have $p^{I_{\chi, v}} = p^{I_{\psi, v}}$. Otherwise $p$ already occurs in $\varphi$ so have $p^{I_{\chi, v}} = p^I$.

	Let $\circ=\to$. Suppose $\mathcal M, v\models\chi$. Then $\mathcal M, v\not\models\varphi$ or $\mathcal M\models\psi$ and therefore $\mathcal M_\chi\not\models\varphi^S_T$ or $\mathcal M_\chi\models\psi^H_T$, i.e. $\mathcal M_\chi\models\chi^S_T$. On the other hand suppose $\mathcal M, v\not\models\chi$. Then $\mathcal M, v\models\varphi$ and $\mathcal M, v\not\models\psi$ and therefore $\mathcal M_\chi, v\models \varphi^S_T$ and $\mathcal M_\chi, v\not\models\psi^S_T$, i.e. $\mathcal M_\chi\not\models\chi^H_T$. Analogous arguments work for $\circ\in\{\wedge, \vee\}$.

	2. $\chi = \forall x\varphi$. Choose $s^{I_\chi}:M^n\to M$ such that $s^{I_\chi}(x_1,\dots, x_n) = m$ if there exists $m\in M$ such that $\mathcal M, v[x_1/t_1\dots x_n/t_n, m/a]\not\models\varphi[a/x]$ for all $v$ and arbitrary otherwise. Since $T$ contains all free variables occurring in $\chi$ this is a well-defined function.

	By induction hypothesis there exists a structure $\mathcal M_{\varphi[a/x]}$ as is the Lemma. Let $p^{I_\chi} = p^{I_{\varphi[a/x]}}$ for all symbols occurring in $\chi$ and $$p^{I_\chi}(x_1\dots x_{i-1}, x_{i+1}\dots x_m) = p^{I_{\varphi[a/x]}}(x_1\dots x_{i-1}, s^{I_\chi}(x_1\dots x_n), x_{i+1}\dots x_m)$$ for all symbols $p$ occurring in $\varphi[a/x]^S_{T\cup a}$ or $\varphi[a/x]^H_{T\cup a}$ but not in $\chi$ where $x_i$ is the argument corresponding to the free variable $a$.

	Suppose $\mathcal M, v\models\chi$. Then for all $m\in M$ $\mathcal M, v[m/a]\models\varphi[a/x]$ and therefore $\mathcal M_{\chi}, v[m/a]\models\varphi[a/x]^S_{M\cup\{a\}}$ and thus $\mathcal M_{\chi}, v\models \forall x\varphi([a/x]^S_{M\cup\{a\}}[x/a])$, i.e. $\mathcal M_\chi,v\models \chi^S$. On the other hand suppose $\mathcal M, v\not\models\chi$. Then there exists $m\in M$ such that $\mathcal M, v[m/a]\not\models\varphi[a/x]$ and therefore $\mathcal M_\chi, v[m/a]\not\models\varphi[a/x]^H_{T\cup\{a\}}$ and so by definition $\mathcal M_\chi, v\not\models\varphi[s(t_1,\dots t_n)/x]^H_T$, i.e. $\mathcal M_\chi, v\not\models(\forall x\varphi)^H_T$.

	3. $\chi = \exists x\varphi$. The argument runs dually to 2. Choose $s^{I_\chi}:M^n\to M$ such that $s^{I_\chi}(x_1,\dots, x_n) = m$ if there exists $m\in M$ such that for all $v$ we have $\mathcal M, v[x_1/t_1\dots x_n/t_n, m/a]\models\varphi[a/x]$ and arbitrary otherwise. Since $T$ contains all free variables occurring in $\chi$ this is a well-defined function.

	By induction hypothesis there exists a structure $\mathcal M_{\varphi[a/x]}$ as is the Lemma. Let $p^{I_\chi} = p^{I_{\varphi[a/x]}}$ for all symbols occurring in $\chi$ and $$p^{I_\chi}(x_1\dots x_{i-1}, x_{i+1}\dots x_m) = p^{I_{\varphi[a/x]}}(x_1\dots x_{i-1}, s^{I_\chi}(x_1\dots x_n), x_{i+1}\dots x_m)$$ for all symbols $p$ occurring in $\varphi[a/x]^S_{T\cup a}$ or $\varphi[a/x]^H_{T\cup a}$ but not in $\chi$ where $x_i$ is the argument corresponding to the free variable $a$.

	Then as in 2 from $\mathcal M, v\models \chi$ follows $\mathcal M_\chi,v\models\chi^S_T$ and from $\mathcal M, v\not\models \chi$ follows $\mathcal M_\chi,v\not\models\chi^H_T$.
\end{proof}

\begin{proof}[Lemma \ref{ap2}]
	Again we proceed by simultaneous induction on the formula height.

	For atoms the claims are clear. We distinguish 5 cases.

	1. $\chi = \varphi\to\psi$. Suppose $\mathcal M, v\models\chi^S_T$, i.e. $\mathcal M, v\not\models\varphi^H_T$ or $\mathcal M, v\models\psi^S_T$. By induction hypothesis $\mathcal M, v\not\models \varphi$ or $\mathcal M, v\models\psi$, i.e. $\mathcal M, v\models \chi$. On the other hand suppose $\mathcal M, v\not\models\chi^H_T$, i.e. $\mathcal M, v\models\varphi^S_T$ and $\mathcal M, v\not\models\varphi^H_T$. Again by induction hypothesis $\mathcal M,v\models\varphi$ and $\mathcal M, v\not\models \varphi$, so $\mathcal M, v\not\models\chi$.

	2. + 3. Conjunctions and Disjunctions are dealt with analogously.

	4. $\chi = \forall x\varphi$.  Suppose $\mathcal M, v\models\chi^S_T$, i.e. for all $m\in M$ we have $\mathcal M, v[m/a]\models \chi[a/x]^S_{T\cup\{a\}}$ and by induction hypothesis $\mathcal M, v[m/a]\models \chi[a/x]$. Then it follows that $\mathcal M, v\models\varphi$. On the other hand suppose $\mathcal M, v\not\models\chi^H_T$, i.e. there exists $m\in M$ such that $\mathcal M, v[m/a]\not\models\chi[a/x]^H_{T\cup \{a\}}$, then by induction hypothesis $\mathcal M, v[m/a]\not\models \chi[a/x]$, i.e. $\mathcal M, v\not\models\forall x\chi$.

	5. $\chi = \exists x\varphi$.  The argument runs dually to 4.
\end{proof}

\end{document}